Network Deliberation: The role of network structure in large-scale, internet-enabled, participatory decision-making

### 2022-08-06 UMich/chapters/ch04-NetDelibExp.tex
Last edits before final draft.
Chapter 4 revisions based on defense feedback.

We focus on deliberation due to fundamental limitations of voting.
Results from social choice theory showt that any ranked-choice voting system is subject to limitations such as the Condorcet Paradox \cite{condorcet_essay_1785} and Arrow's Impossibility Theorem \cite{arrow_social_2012}.
Deliberation offers one way to avoid these limitations by changing individual preferences \cite{ackerman_deliberation_2002}.
The success of deliberation at generating consensus has been attributed to its ability to identify and resolve conflict \cite{gentry_consensus_1982}.
However, deliberation also carries risks, such as the posibility of pushing individuals towards extreme and polarized preferences \cite{schkade_what_2007}.

In any sufficiently large group, it becomes unfeasible for all members to interact with each other.
Understanding collaboration in large groups thus requires attention to the network structure of who interacts with whom.
The field of social learning has studied both simple estimation tasks \cite{degroot_reaching_1974, golub_naive_2010} and more complex rugged-landscape optimization tasks \cite{lazer_network_2007, barkoczi_social_2016}.
Observational studies have been used to analyze the network structures of online collaborations \cite{gonzalez-bailon_networked_2016, platt_network_2018}.
Lab and field studies have also been used to study real-world collaborations on a small scale \cite{salehi_hive_2018, kearns_experiments_2012}.
As discussed in Chapter \ref{chap:abm}, the successful large-scale collaborations enabled by the internet often exhibit network sructures characterized by small, interlocking groups, reminiscent of interlocking directorates \cite{levine_study_1979} and interlocking publics \cite{habermas_structural_1991}.
In Chapter \ref{chap:abm}, we describe agent-based simulations showing that such network deliberation can improve performance on complex tasks when individuals exhibit strong social influence.dsd
Here, we expand on existing research by performing a controlled online experimental evaluation of network deliberation.
We use periodic polling to track preference evolution, allowing causal inference in a real-world setting.

While many studies have examined the role of network structure in collaboration, there has so far been little attention to the interlocking pod structure of network deliberation.
The closest we are aware of is the study of network rotation by Salehi \& Bernstein \cite{salehi_hive_2018}.
The present experiment goes beyond existing work primarily by tracking preferences throughout the experiment and by the use of a conventional single-group control condition.
We also describe multiple network deliberation topologies for achieving efficient or inefficient networks.
For the present study, we focus on an efficient topology.

@@

\section{Theory}
\label{sec:theory}
\subsection{Social Choice Theory}
Our analysis of preferences is foudned in the formalism of social choice theory \cite{arrow_social_2012, gaertner_primer_2009}.
In social choice theory, individual's {\em preferences} are represented by a ranked ordering of {\em alternatives} (e.g., proposals, candidates).
The set of preferences for all individuals in a group is called the {\em preference profile}.
A {\em social welfare function} can be applied to the profile to generate a {\em social preference}: an ordering of alternatives representing the preference of the group as a whole.
A {\em social choice function} represents a voting system: it selects a subset of winners from the available alternatives.

@@

\subsubsection{Weighted Crossing Distance}
We propose the following weighted distance metric for preference rankings.
Given a set of $M$ alternatives, there are $M-1$ ways to divide an ordering into two nonempty sets of adjacent alternatives, i.e., a high set and a low set.
For a pair of preferences $p, q$, we define a length $M - 1$ {\em crossing vector} $v_m(p, q)$ to be equal to the number of alternatives in the highest $m$ positions of $p$ and not in the highest $m$ positions of $q$ (it is easily shown that this measure is symmetric in $p$ and $q$).
Informally, $v_m$ is the number of alternatives that cross place $m$.
Each element of $v_m$ is an integer in the range $[0, v_m^{\max}]$ dependent on the index:
\begin{equation}
    v_m^{\max} =
    \begin{cases}
    m & m \leq (M - 1) / 2 \\
    M - m & (M - 1) / 2 < m \leq M - 1.
    \end{cases}
\end{equation}
We construct the weighted crossing distance by taking a weighted sum over this vector.
There are many potential choices for weights.
We choose the weight of place $m$ to be equal to the highest total of all lower ranked places plus one.
Thus the weight at index $i = M - m$ (noting that low rank corresponds to high $m$) is given by:
\begin{equation}
    w_i = 1 + \sum_{k=1}^{i-1} w_k v_k^{\max}.
\end{equation}

@@

We are particularly interested in the role of social influence in preference evolution and require a formalism to quantify and analzye social influence.
If two individuals preferences become closer, can we say whether one influenced the other? And if so, can we say which influence which?
We cannot answer either question with certainty from preferences alone, but we can construct a reasonable proxy based on some simplifying assumptions.
Namely, we will construct measures of social influence based on the assumptions that:
\begin{itemize}
    \item The magnitude of an individual's change in preference increases with their susceptibility with social influence;
    \item The magnitude of an individual's change in preference increases with the magnitude of the distance between their preference and the influencing preference.
\end{itemize}
Thus, if the distance between the preferences of individuals $A$ and $B$ decreases, and the preference of $B$ changed more than the preference of $A$, we assume that individual $A$ influenced individual $B$.
We now describe two formal measures that use these assumptions to quantify how susceptible to influence and how potential influential an individual is.

@@

We wish to quantify whether a change in agreement/disagreement is due to individuals converging towards initially popular alternatives, or to the the group being persuaded towards initially unpopular alternatives.
In the former case, we would expect the majority of individuals to change their preferences by a small amount, while a smaller number of individuals holding unpopular opinions make larger shifts towards popular preferences.
In the latter case, we would expect the opposite:
some individuals with unpopular preferences hold closely to those opinions, while a large number of individuals move away from the previous popular preferences.
So we are interested not just in the total change in agreement/disagreement,
but also in how that change is distributed across the individuals in the group.

To apportion the change in agreement between individuals $p$ and $q$,
we make the assumption that an individual's contribution to the change in agreement is proportional to the change in their own preference.
Specifically, we define the conformity $c(p,q)$ of individual $p$ toward individual $q$ as:
\begin{eqnarray}
    d_i &=& d(p_i, q_i) \\
    d_f &=& d(p_f, q_f) \\
    |p| &=& d(p_i, p_f) \\
    c(p,q) &=& -(d_f - d_i)\frac{|p|}{|p| + |q|}.
\end{eqnarray}
Note that the change in agreement between two individuals can be written in terms of the conformity:
\begin{equation}
    -(d_f - d_i) = c(p, q) + c(q, p).
\end{equation}
Similarly, the total change in agreement (\ref{eqn:Delta}) can be rewritten as a sum of conformity over all pairs of individuals:
\begin{equation}
    \Delta =
    \sum_p \sum_{q \neq p} c(p, q).
\end{equation}
By performing only one of the sums above, the change in agreement $\Delta$ can be written as a sum of terms, each corresponding to one individual.
The sum can be performed over either the first or second argument of $c$.
If the sum is performed over the second argument of $c$, the result is the mean conformity of a particular individual $p$ toward all others:
\begin{equation}
    C(p) = \sum_q c(p, q).
\end{equation}
Alternatively, if the sum is performed over the first argument of $c$, the result is the mean conformity of all others toward $p$, which we call the hipness $H(p)$:
\begin{equation}
    H(p) = \sum_q c(q, p).
\end{equation}
Both $C(p)$ and $H(p)$ are equal to $\Delta$ when summed over all individuals, but they describe conceptually different attributes of an individual.
The conformity quantifies how much an individual has increased agreement by changing their preferences to match others.
The hipness quantifies how much an individual has increased agreement due to holding an unchanging preference near to others' final preferences.

@@

\begin{figure}
\label{fig:kcon}
\center
\includegraphics[width=5in]{chapters/figures/NetDelibExp/fig-kcon}
\caption{The k-connectivity of long-path and random-pod network deliberation
networks for $T=3$ stages and a pod size of $M=5$.}
\end{figure}

@@

\begin{table}
\center
\label{tab:prompts}
\begin{tabular}{|p{0.5in}|p{4.5in}|}
\hline
Stage & Prompt \\
\hline
1 &
Which proposals do you prefer, and why?
If there are disagreements, try asking questions to understand the perspective
of others and to understand the causes of the disagreement.
\\
\hline
2 & In the previous round of discussion, what opinions and reasoning did you
observe in your group?
How much agreement was there?
Were there any disagreements or conflicts?
If so, what were the sources of conflict, and how might they be resolved?
\\
\hline
3 &
What seem to be the most popular opinions?
Do you agree with them?
Has your opinion changed over the course of the discussion?
\\
\hline
\end{tabular}
\caption{Prompts shown to participants during each round of deliberation.}
\end{table}

@@

Participants were assigned to one of two conditions at the time of enrollment.
Assignments were uniformly at random.
Stratified sampling would be preferable in order to ensure even demographic distributions, but was not feasible due to software and time constraints.
However, we manually confirmed that each condition had comparable demographics.
The two conditions were:
\begin{description}
\item[Control]{
Participants in the control condition engaged in conventional deliberation in a single large group.
Posts and comments by participants in the control condition were visible
to all others in the control condition for the duration of the deliberation.}
\item[Random-Pod (Network Deliberation)]{
Participants in this condition were divided into small pods ($\leq 5$ participants).
Posts and comments made by participants in this condition were only visible to others in their current pod.
Participants in this condition were assigned to a new pod at the beginning of each round of deliberation using the random-pod assignment method, producing a structurally efficient communication network.}
\end{description}

@@

\begin{description}
\item[Prop. 1]{Open-ended with partial credit (2 questions, 5 points each) - Ch. 1-2: Intro to networks.}
\item[Prop. 2]{Open-ended with partial credit (2 questions, 5 points each) - Ch. 3: Bridges, clustering, triadic closure, strong and weak ties, and the Strong Triadic Closure Property.}
\item[Prop. 3]{Multiple choice with no partial credit (5 questions, 2 points each) - Ch. 4: Homophily, selection, social influence, and social-affiliation networks.}
\item[Prop. 4]{Multiple choice with no partial credit (5 questions, 2 points each) - Ch. 5: Signed networks and balance theory.}
\item[Prop. 5]{True/false with no partial credit (10 questions, 1 point each) - Ch. 6: Game theory, best responses, and Dominant Strategies, Nash equilibrium.}
\item[Prop. 6]{True/false with no partial credit (10 questions, 1 point each) - Ch. 9:  Auctions.}
\end{description}
\end{displayquote}

@@

\begin{figure}
\label{fig:reply}
\center
\frame{\includegraphics[width=5in]{chapters/figures/NetDelibExp/fig-reply.png}}
\caption{Participants can perform standard forum actions, such as posting and commenting.}
\end{figure}

\begin{figure}
\label{fig:vote}
\center
\frame{\includegraphics[width=5in]{chapters/figures/NetDelibExp/fig-vote.png}}
\caption{Participants are presented with a ranked-choice voting screen between rounds of deliberation.}
\end{figure}

\begin{figure}
\label{fig:alias}
\center
\frame{\includegraphics[width=3in]{chapters/figures/NetDelibExp/fig-alias.png}}
\caption{Participants are assigned an alias for the duration of the experiment.}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=3in]{chapters/figures/NetDelibExp/fig-counts.pdf}
    \caption{Number of first-choice votes for each alternative over the course of the deliberation.}
    \label{fig:firstplace}
\end{figure}

Over the course of the deliberation, the preference profiles of the control and random-pod groups show notable differences.
In the control group, Prop. 1 becomes the Condorcet winner after round 1 and wins across all other voting methods.
Prop 1. remains the winner in the control group throughout the rest of the deliberation, with one exception: Prop. 1 and Prop. 2 tie for plurality voting in round 2.
In the random-pod group, Prop. 2 remains the winner across all voting methods throughout the entire deliberation.
We compare participants' preferences before and after the deliberation in Figure \ref{fig:countmatrix}.
The most striking feature is that a number of participants (4) in the control group who initially ranked Prop. 2 highest then went on to rank Prop 1. the highest in the final poll.
These participants are a significant driver of the switch from Prop 2. to Prop 1. in the control group.
While the random-pod group does show some participants switching away from Prop. 2, those participants shift to the less popular Props. 3 and 4 rather than Prop 1.
In the sections that follow, we will further analyze the evolution of participant preferneces and the potential mechanisms driving them.

@@

Here we describe the trends we observe in participants' deliberation activity and the potential impact on participant preferences.
Overall, the nubmer of comments per participant shows a heavy-tailed distribution (Figure \ref{fig:commentcounts}) with a mode of one comment per participant.
The random-pod group shows less of a heavy-tail, with more participants posting a single comment per round and a smaller maximum count (2--3 comments vs. 3--4 for control).
The random-pod group thus shows a more equal division of activity across participants.

@@

Figure \ref{fig:commentrounds} shows the comment counts for each round as well as the Shannon entropy of participant comment counts.
The random-pod group shows a lower total number of comments and a higher entropy, again showing a more equal division of activity across participants.

@@

Our results show that in our experiment, the random-pod group 
experienced a small but robust increase in agreement,
wile the control group experienced a sharp increase followed by an even sharper decrease.
These results support the hypothesis H1 that network deliberation contributes to increased agreement relative to conventional single-group deliberation.
However, this effect is due more to a decrease in agreement
within the control group than an increase in the random-pod group.
After an initial increase in agreement, both groups showed an unexpected ``rebound'' effect.
In the control group, this rebound resulted in a net reduction in agreement, suggesting the presence of negative social influence.

@@

In both the control and random-pod groups, we see round 1 deliberation clearly favoring one alternative, which then gains a number of first-choice votes, suggesting social influence.
This social influence is further evidenced by the increase in agreement observed for both groups over the course round 1 (Figure \ref{fig:agreement}).

@@

Combining the above insights from conformity and hipness measures suggests the following.
The negative shift in agreement in the control group is due
to a majority of participants shifting their preferences away (negative conformity) from a samll number of stubborn participants (highly-negative hipness) without converging on an alternative (lack of positive conformity).
Furthermore, this shift occurs after round 2, following an initial stage of high-conformity.
The random-pod participants also exhibit stubborn participants, but fewer and with smaller magnitude,
and the the majority of participants converge toward each other's preferences (positive conformity).

@@

We hypothesized (H1) that network deliberation tends to reach a higher level of agreement than conventional single-group deliberation.
The observed agreement calculated from polls throughout the experiment experiment supports this hypothesis (Section \ref{sec:res-agreement}).
However, our hypothesis was based on the assumption that network deliberation would more effectively identify and resolve disagreements,
leading us to expect a large increase increase in agreement in the random-pod group.
Instead, we observe a small increase of agreement within the random-pod group and a large decrease within the control group.
So while we do see more of an improvement in agreement under network deliberation, our observations are more consistent with a protective effect against a disagreement-producing mechanism, than with the originally hypothesized agreement-producing mechanism.

>>>>

We focus on deliberation due to fundamental limitations of voting.
Results from social choice theory show that any ranked-choice voting system is subject to limitations such as the Condorcet Paradox \cite{condorcet_essay_1785} and Arrow's Impossibility Theorem \cite{arrow_social_2012}.
Deliberation offers one way to avoid these limitations by changing individual preferences \cite{ackerman_deliberation_2002}.
The success of deliberation at generating consensus has been attributed to its ability to identify and resolve conflict \cite{gentry_consensus_1982}.
However, deliberation also carries risks, such as the posibility of pushing individuals towards extreme and polarized preferences \cite{schkade_what_2007}.

In any sufficiently large group, it becomes unfeasible for all members to interact with each other.
Understanding collaboration in large groups thus requires attention to the network structure of who interacts with whom.
The field of social learning has studied both simple estimation tasks \cite{degroot_reaching_1974, golub_naive_2010} and more complex rugged-landscape optimization tasks \cite{lazer_network_2007, barkoczi_social_2016}.
Observational studies have been used to analyze the network structures of online collaborations \cite{gonzalez-bailon_networked_2016, platt_network_2018}.
Lab and field studies have also been used to study real-world collaborations on a small scale \cite{salehi_hive_2018, kearns_experiments_2012}.
As discussed in Chapter \ref{chap:abm}, the successful large-scale collaborations enabled by the internet often exhibit network sructures characterized by small, interlocking groups, reminiscent of interlocking directorates \cite{levine_study_1979} and interlocking publics \cite{habermas_structural_1991}.
In Chapter \ref{chap:abm}, we describe agent-based simulations showing that such network deliberation can improve performance on complex tasks when individuals exhibit strong social influence.
Here, we expand on existing research by performing a controlled online experimental evaluation of network deliberation.
We use periodic polling to track preference evolution, allowing causal inference in a real-world setting.

While many studies have examined the role of network structure in collaboration, there has so far been little attention to the interlocking pod structure of network deliberation.
The closest we are aware of is the study of network rotation by Salehi \& Bernstein \cite{salehi_hive_2018}.
The present experiment goes beyond existing work in several ways.
Our primarity innovation is the tracking of preferences throughout the experiment,
which allows the quantification of consensus and social influence.
We also include a large single-group control condition.
Through the use of a pseudonymous online platform, we isolate deliberative text and network structure from other social cues that
might result in social influence, e.g., tone of voice, visual appearance.
While these factors undoubtedly play an important role in the social dynamics of deliberation,
we are focused specifically on the role of communication network structure.
We also describe multiple network deliberation topologies for achieving efficient or inefficient networks.
For the present study, we focus on an efficient topology.
By tracking preference evolution, limiting confounding factors, and using a controlled experimental design,
we can examine the dynamics and causality of the relationship between network structure and preference evolution.

Observational and numerical findings have shown that successful large-scale collaborations often exhibit a structure
of small, interlocking pods: {\em network deliberation}.
This pod structure can be seen in a range of collaborations, including the
articles of Wikipedia \cite{platt_network_2018},
the projects and modules of the free/open-source software community \cite{benkler_coases_2002},
and the nodes of Occupy Sandy \cite{brugh_combining_2019}.
Numerical simulations (Chapter \ref{chap:abm}) demonstrate the effectiveness of the network
deliberation structure in the presence of social influence.

@@

\section{Theory}
\label{sec:theory}
\subsection{Social Choice Theory}
Our analysis of preferences is founded in the formalism of social choice theory \cite{arrow_social_2012, gaertner_primer_2009}.
In social choice theory, individual's {\em preferences} are represented by a ranked ordering of {\em alternatives} (e.g., proposals, candidates).
The set of preferences for all individuals in a group is called the {\em preference profile}.
A {\em social welfare function} can be applied to the profile to generate a {\em social preference}: an ordering of alternatives representing the preference of the group as a whole.
A {\em social choice function} represents a voting system: it selects a subset of winners from the available alternatives.

@@

\subsubsection{Weighted Crossing Distance}
We propose the following weighted distance metric for preference rankings.
Given a set of $M$ alternatives, there are $M-1$ ways to divide an ordering into two nonempty sets of adjacent alternatives, i.e., a high set and a low set.
For a pair of preferences $p, q$, we define a length $M - 1$ {\em crossing vector} $v_m(p, q)$ to be equal to the number of alternatives in the highest $m$ positions of $p$ and not in the highest $m$ positions of $q$ (it is easily shown that this measure is symmetric in $p$ and $q$).
Informally, $v_m$ is the number of alternatives that cross place $m$.
Each element of $v_m$ is an integer in the range $[0, v_m^{\max}]$ dependent on the index:
\begin{equation}
\label{eqn:distance-max}
    v_m^{\max} =
    \begin{cases}
    m & m \leq (M - 1) / 2 \\
    M - m & (M - 1) / 2 < m \leq M - 1.
    \end{cases}
\end{equation}
We construct the weighted crossing distance by taking a weighted sum over this vector.
There are many potential choices for weights.
We choose the weight of place $m$ to be equal to the highest total of all lower ranked places plus one.
Thus the weight at index $i = M - m$ (noting that low rank corresponds to high $m$) is given by:
\begin{equation}
\label{eqn:weight}
    w_i = 1 + \sum_{k=1}^{i-1} w_k v_k^{\max}.
\end{equation}

For example, consider the following two preference orderings (in least-to-most preferred order):
\begin{eqnarray}
p &=& (C, B, E, F, A, D) \\
q &=& (F, E, B, C, A, D).
\end{eqnarray}
Applying Eq. \ref{eqn:distance-max} and \ref{eqn:weight} 
gives the maximum place-values $v^{\text{max}}$ and place weights $w_i$:
\begin{eqnarray}
v^{\text{max}} &=& (1, 2, 3, 2, 1) \\
w &=& (1, 2, 6, 24, 72).
\end{eqnarray}
Partitioning both vectors at index 3 gives:
\begin{eqnarray}
p &\mapsto& {C, B, E}, {F, A, D} \\
q &\mapsto& {F, E, B}, {C, A, D}.
\end{eqnarray}
Noting that these partitions differ by a single swap ($F \rightleftarrows C$)
the corresponding element of the crossing vector $v_3$ is $1$.
Applying the same procedure to each index gives the full crossing vector:
\begin{eqnarray}
v(p,q) &=& (1, 2, 1, 0, 0).
\end{eqnarray}
Taking the inner product with the weight vector gives the total crossing distance:
\begin{eqnarray}
d(p, q) &=& v(p,q) \cdot w \\
&=& 11.
\end{eqnarray}
Normalizing by the maximum crossing distance $143$ gives a normalized distance of approximately $0.08$.
For comparison, the normalized Kendall tau distance bewtween these two preferences is $0.40$,
much larger despite agreeing on the highest two choices.

@@

Deliberation is fundamentally a social influence process in which
individuals exchange statements with their peers in an effort to influence their preferences.
We will thus benefit from a formalism to quantify and analzye social influence.
If two individuals preferences become closer, can we say whether one influenced the other? And if so, can we say which influence which?
We cannot answer either question with certainty from preferences alone, but we can construct a reasonable proxy based on some simplifying assumptions.
Namely, we will construct measures of social influence based on the assumptions that:
\begin{itemize}
    \item The magnitude of an individual's change in preference increases with their susceptibility to social influence;
    \item The magnitude of an individual's change in preference increases with the magnitude of the distance between their preference and the influencing preference.
\end{itemize}
The distance between two individuals' preferences is symmetric between the two individuals,
so the latter of the above conditions does not speak to who influences whom.
The converse of the former condition however suggests that the greater the change in an individual's preference,
the more they have been influenced by other.
Thus, if the distance between the preferences of individuals $A$ and $B$ decreases, and the preference of $B$ changed more than the preference of $A$, we assume that individual $A$ influenced individual $B$.
On the other hand, if the distance increases, we refer to this as negative social influence, suggesting one individual has influenced another to adopt an opposing preference.
We now describe two formal measures that use these assumptions to quantify how susceptible to influence and how potential influential an individual is.

@@

Measuring only the change in agreement leaves some ambiguities.
For instance, changes in agreement could simply be due to individuals converging towards initially popular preferences.
Alternatively initially unpopular preferences might overtake others in popularity,
which suggests unequal social influence, and might indicate the presence of an information cascade.
We wish to quantify whether a change in agreement/disagreement is due to individuals converging towards initially popular alternatives, or to the the group being persuaded towards initially unpopular alternatives.
In the former case, we would expect the majority of individuals to change their preferences by a small amount, while a smaller number of individuals holding unpopular opinions make larger shifts towards popular preferences.
In the latter case, we would expect the opposite:
some individuals with unpopular preferences hold closely to those opinions, while a large number of individuals move away from the previous popular preferences.
So we are interested not just in the total change in agreement/disagreement,
but also in how that change is distributed across the individuals in the group.

We divide the change in agreement between individuals $p$ and $q$ into two parts:
that due to $p$'s influence on $q$ and vice versa. 
To apportion the change in agreement between individuals $p$ and $q$,
we make the assumption that the contribution due to $q$'s influence on $p$'s preferences is
proportional to the change in $p$'s preference.
Specifically, we define the conformity $c(p,q)$ of individual $p$ toward individual $q$ as:
\begin{eqnarray}
    d_i &=& d(p_i, q_i) \\
    d_f &=& d(p_f, q_f) \\
    |p| &=& d(p_i, p_f) \\
    c(p,q) &=& -(d_f - d_i)\frac{|p|}{|p| + |q|}.
\end{eqnarray}
These values are illustrated schematically in Figure \ref{fig:conform-hip-diagram}.
Note that the change in agreement between two individuals can be written in terms of the conformity:
\begin{equation}
    -(d_f - d_i) = c(p, q) + c(q, p).
\end{equation}
Similarly, the total change in agreement (\ref{eqn:Delta}) can be rewritten as a sum of conformity over all pairs of individuals:
\begin{equation}
    \Delta =
    \sum_p \sum_{q \neq p} c(p, q).
\end{equation}
By performing only one of the sums above, the change in agreement $\Delta$ can be written as a sum of terms, each corresponding to one individual.
The sum can be performed over either the first or second argument of $c$.
If the sum is performed over the second argument of $c$, the result is the mean conformity of a particular individual $p$ toward all others:
\begin{equation}
    C(p) = \sum_q c(p, q).
\end{equation}
Alternatively, if the sum is performed over the first argument of $c$, the result is the mean conformity of all others toward $p$, which we call the hipness $H(p)$:
\begin{equation}
    H(p) = \sum_q c(q, p).
\end{equation}
Note that a high hipness does not imply that an individual has influenced others to adopt similar preferences,
only that others have adopted such preferences.
Both $C(p)$ and $H(p)$ are equal to $\Delta$ when summed over all individuals, but they describe conceptually different attributes of an individual.
The conformity quantifies how much an individual has increased agreement by changing their preferences to match others.
The hipness quantifies how much an individual has increased agreement due to holding an unchanging preference near to others' final preferences.

\begin{figure}
\centering
\includegraphics[width=3in]{chapters/figures/NetDelibExp/fig-conform-hip-diagram.pdf}
\caption{Schematic illustration of the quantities used in calculating the social influence
between $p$ and $q$. Initial preferences are shown as $p_i$ and $q_i$ with final preferences
shown as $p_f$ and $q_f$. The change in agreement from initial to final is primarily due to
$p$'s movement, so we consider $p$ to be high in conformity, while $q$ is high in hipness.}
\label{fig:conform-hip-diagram}
\end{figure}

@@

\begin{figure}
\center
\includegraphics[width=5in]{chapters/figures/NetDelibExp/fig-kcon}
\caption{The k-connectivity of long-path and random-pod network deliberation
networks for $T=3$ stages and a pod size of $M=5$.}
\label{fig:kcon}
\end{figure}

@@

\begin{table}
\center
\begin{tabular}{|p{0.5in}|p{4.5in}|}
\hline
Stage & Prompt \\
\hline
1 &
Which proposals do you prefer, and why?
If there are disagreements, try asking questions to understand the perspective
of others and to understand the causes of the disagreement.
\\
\hline
2 & In the previous round of discussion, what opinions and reasoning did you
observe in your group?
How much agreement was there?
Were there any disagreements or conflicts?
If so, what were the sources of conflict, and how might they be resolved?
\\
\hline
3 &
What seem to be the most popular opinions?
Do you agree with them?
Has your opinion changed over the course of the discussion?
\\
\hline
\end{tabular}
\caption{Prompts shown to participants during each round of deliberation.}
\label{tab:prompts}
\end{table}

@@

Participants were assigned to one of two conditions at the time of enrollment.
Assignments were uniformly at random.
We manually confirmed that each condition had comparable demographics.
The two conditions were:
\begin{description}
\item[Control]{
Participants in the control condition engaged in conventional deliberation in a single large group.
Posts and comments by participants in the control condition were visible
to all others in the control condition for the duration of the deliberation.}
\item[Random-Pod (Network Deliberation)]{
Participants in this condition were divided into small pods ($\leq 5$ participants).
Posts and comments made by participants in this condition were only visible to others in their current pod.
Participants in this condition were assigned to a new pod at the beginning of each round of deliberation using the random-pod assignment method, producing a structurally efficient communication network.}
\end{description}

\subsubsection{Ethical Considerations}
This study was reviewed by the University of Michigan's IRB and classified as exempt.
To protect participant privacy, all participants were assigned randomly-generated pseudonyms.
The keys connecting pseudonyms to identifiable information were stored securely and only available
to the researchers.
Participants were provided with an email address to report any harassment, which was continuously
monitored throughout the experiment.
The deliberations were also regularly reviewed for any disruptive, hateful, or harassing activity.

@@

\begin{description}
\item[Prop. 1]{Open-ended, partial credit (2 questions, 5 points each) - Ch. 1-2}
\item[Prop. 2]{Open-ended, partial credit (2 questions, 5 points each) - Ch. 3}
\item[Prop. 3]{Multiple choice, no partial credit (5 questions, 2 points each) - Ch. 4}
\item[Prop. 4]{Multiple choice, no partial credit (5 questions, 2 points each) - Ch. 5}
\item[Prop. 5]{True/false, no partial credit (10 questions, 1 point each) - Ch. 6}
\item[Prop. 6]{True/false, no partial credit (10 questions, 1 point each) - Ch. 9}
\end{description}
\end{displayquote}

@@

\begin{figure}
\center
\frame{\includegraphics[width=5in]{chapters/figures/NetDelibExp/fig-reply.png}}
\caption{Participants can perform standard forum actions, such as posting and commenting.}
\label{fig:reply}
\end{figure}

\begin{figure}
\center
\frame{\includegraphics[width=5in]{chapters/figures/NetDelibExp/fig-vote.png}}
\caption{Participants are presented with a ranked-choice voting screen between rounds of deliberation.}
\label{fig:vote}
\end{figure}

\begin{figure}
\center
\frame{\includegraphics[width=3in]{chapters/figures/NetDelibExp/fig-alias.png}}
\caption{Participants are assigned an alias for the duration of the experiment.}
\label{fig:alias}
\end{figure}

@@

\begin{figure}
    \centering
    \raisebox{-\height}{\includegraphics[width=3in]{chapters/figures/NetDelibExp/fig-counts.pdf}}
    \raisebox{-\height}{\includegraphics[width=1in]{chapters/figures/NetDelibExp/fig-counts-legend.pdf}}
    \caption{Number of first-choice votes for each alternative over the course of the deliberation.}
    \label{fig:firstplace}
\end{figure}

Over the course of the deliberation, the preference profiles of the control and random-pod groups show notable differences.
In the control group, Prop. 1 becomes the Condorcet winner after round 1 and wins across all other voting methods.
Prop 1. remains the winner in the control group throughout the rest of the deliberation, with one exception: Prop. 1 and Prop. 2 tie for plurality voting in round 2.
In the random-pod group, Prop. 2 remains the winner across all voting methods throughout the entire deliberation.
We compare participants' preferences before and after the deliberation in Figure \ref{fig:countmatrix}.
The most striking feature is that a number of participants (4) in the control group who initially ranked Prop. 2 highest then went on to rank Prop 1. the highest in the final poll.
These participants are a significant driver of the switch from Prop 2. to Prop 1. in the control group.
While the random-pod group does show some participants switching away from Prop. 2, those participants shift to the less popular Props. 3 and 4 rather than Prop 1.
In the sections that follow, we will further analyze the evolution of participant preferences and the potential mechanisms driving them.

@@

Here we describe the trends we observe in participants' deliberation activity and the potential impact on participant preferences.
Overall, the number of comments per participant shows a heavy-tailed distribution (Figure \ref{fig:commentcounts}) with a mode of one comment per participant.
The random-pod group shows less of a heavy-tail, with more participants posting a single comment per round and a smaller maximum count (2--3 comments vs. 3--4 for control).
The random-pod group thus shows a more equal division of activity across participants.

@@

Figure \ref{fig:commentrounds} shows the comment counts for each round as well as the Shannon entropy of the distribution of comment counts over participants.
The random-pod group shows a lower total number of comments and a higher entropy, again showing a more equal division of activity across participants.

@@

Our results show that in our experiment, the random-pod group 
experienced a small but robust increase in agreement,
wile the control group experienced a sharp increase followed by an even sharper decrease.
These results support the hypothesis H1 that network deliberation contributes to increased agreement relative to conventional single-group deliberation.
However, this effect is due more to a decrease in agreement
within the control group than an increase in the random-pod group.
After an initial increase in agreement, both groups showed an unexpected ``rebound'' effect.
In the control group, this rebound resulted in a net reduction in agreement, suggesting the presence of negative social influence.
In other words, participants altered their preferences to more strongly oppose those they interacted with.

@@

In both the control and random-pod groups, we see round 1 deliberation clearly favoring one alternative, which then gains a number of first-choice votes, suggesting social influence.
This social influence is further evidenced by the increase in agreement observed for both groups over the course of round 1 (Figure \ref{fig:agreement}).

@@

Combining the above insights from conformity and hipness measures suggests the following.
The negative shift in agreement in the control group is due
to a majority of participants shifting their preferences away (negative conformity) from a small number of stubborn participants (highly-negative hipness) without converging on an alternative (lack of positive conformity).
Furthermore, this shift occurs after round 2, following an initial stage of high-conformity.
The random-pod participants also exhibit stubborn participants, but fewer and with smaller magnitude,
and the majority of participants converge toward each other's preferences (positive conformity).

@@

We hypothesized (H1) that network deliberation tends to reach a higher level of agreement than conventional single-group deliberation.
The observed agreement calculated from polls throughout the experiment experiment supports this hypothesis (Section \ref{sec:res-agreement}).
However, our hypothesis was based on the assumption that network deliberation would more effectively identify and resolve disagreements,
leading us to expect a large increase in agreement in the random-pod group.
Instead, we observe a small increase of agreement within the random-pod group and a large decrease within the control group.
So while we do see more of an improvement in agreement under network deliberation, our observations are more consistent with a protective effect against a disagreement-producing mechanism, than with the originally hypothesized agreement-producing mechanism.
